Utilizing Convolution for Obstacle Avoidance during Robotic Manipulator Movement

### 2023-07-27 Articles/ERK/erk.tex
ERK .
.
ERK references, eq descriptions

\usepackage[utf8]{inputenc}
\usepackage{erk}
\usepackage{times}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage[top=22.5mm, bottom=22.5mm, left=22.5mm, right=22.5mm]{geometry}

@@

\todocomment{Dodaj opis poglavij.}

@@

@@

\section{Methodology }

@@

The Jacobian matrix is a key element in the control of robotic manipulators. It can also be used for computing the inverse kinematics problem. However, for a 7-DOF manipulator, the Jacobian matrix is not square and its inverse does not exist in a traditional sense. In such cases, we resort to the pseudo-inverse of the Jacobian, calculated using the Moore-Penrose method. The Moore-Penrose pseudo-inverse, denoted as $J^+$, provides an approximation of the inverse for non-square matrices.

Damping, represented by the term $k_{damp}$ in the Moore-Penrose pseudo-inverse equation, is used to introduce stability and prevent potential numerical issues when computing the pseudo-inverse of the Jacobian for non-square matrices. It acts as a regularization term to ensure a well-conditioned solution.

@@

The null space of the Jacobian corresponds to those joint motions which produce no movement in the end-effector. The primary task for our robot is the desired end-effector movement. The secondary task of obstacle avoidance in our research takes advantage of this null-space motion~\cite{lajpah_Petri_2012}. Our method applies a repulsive velocity from detected obstacles in the environment. We calculate this velocities at certain points that are positioned on the manipulator arm segments. The robot arm is thus able to carry out the primary task while simultaneously avoiding obstacles by appropriately utilizing its redundant DOF.

\todocomment{cite Tadej?}

@@

\todocomment{check if "exact equation" gives better results}

\todocomment{add references to equations in text}

@@

By employing the Hadamard (element-wise) product between the segment of the obstacle grid \(A_d\) and the corresponding 3D convolutional kernel \(K_d\) for direction \(d\), we derive the resultant matrix \(C_d\), which can be expressed as:

@@

Subsequently, the avoidance velocity in the Cartesian coordinate system is obtained by summing all the values in the resulting matrix, which can be represented as:

\begin{equation}
v = \sum_{i}\sum_{j}\sum_{k} a_{ijk} 
\label{summation}
\end{equation}

@@

Our convolutional kernels are essentially discrete differential operators in 3D space, and their design reflects this. These kernels bear a resemblance to the Sobel kernels, which are widely used in computer vision for edge detection and approximation of spatial derivatives. Each kernel has a primary axis aligned with the direction in which the avoidance velocity is being calculated, with the other two axes orthogonal to it. The kernel values along the primary axis exhibit a symmetric pattern: the highest value lies in the center of the kernel, with the absolute values decreasing uniformly on either side, with negative polarity on one side. This decrease simulates the derivative-like behavior of the kernel.

\todocomment{add references to images in text}

@@

Contrary to the design of a classical Sobel operator, our custom kernels don't contain zeroes along the primary axis. Instead, we assign the maximum kernel values along this line. This choice is specifically designed to meet our goal of obstacle avoidance. By doing so, when our points of interest on the robot manipulator coincide with an obstacle - effectively being 'right on top' of the obstacle in the voxel grid - the convolution operation yields a maximal repulsive velocity. This encourages the robot to move away from the obstacle and ensures safer navigation in the presence of potential collisions.

@@

\todocomment{while not the most important fix, currently values in matrix don't follow values used in experiments}

In addition to the kernel values along the primary axis, the values along the orthogonal axes mimic the same pattern. These are derived from the primary axis values, decrease proportionally with the radial distance from the primary axis, and crucially, they maintain the same sign as dictated by the derivative along the primary axis. This symmetry in both magnitude and sign around the main axis is a fundamental characteristic of our kernels.

@@

The dimensions of the kernel, both along the primary axis and orthogonal to it, play crucial roles in determining the effective detection range and the nature of the repulsive avoidance velocity.

@@

Additionally, we can augment our velocity component with repulsive fields, which are calculated using the convolutional kernel technique. This enables us to guide the end-effector around obstacles smoothly, ensuring safe and collision-free navigation while dynamically adapting to the environment. By integrating both the attractive and repulsive field contributions, our motion planning approach ensures efficient and effective trajectory generation, enabling the robotic manipulator to reach its desired goal while avoiding potential obstacles in the workspace.

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=1.0\linewidth]{Repulsive-Velocities-Along-the-Y-Axis-wide.eps}
	\caption{Plot depicting avoidance velocities (x, y, z) as calculated by our kernel along the y-axis, showing the robot's adaptive movement away from the wall obstacle.}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=1.0\linewidth]{Y-Positions.eps}
	\caption{Comparative visualization of the y-position of the robot's fourth joint with and without the secondary avoidance task, highlighting the impact of the repulsive force.}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=1.0\linewidth]{7-DOF_Robot_Positions_Standard_vs_Avoidance.eps}
	\caption{3D model displaying the final position of the robot, contrasting the scenarios with and without the implementation of joint 4 obstacle avoidance.}
\end{figure}

@@

The end-effector's motion was controlled using the attractive field method, serving as the primary task. This was complemented by the repulsive field method implemented as a secondary task, with the specific aim of maneuvering a point on the fourth joint of the robot as far away from the obstacle - in this case, a simple wall - as possible.

@@

Figure 3 illustrates the avoidance velocities in the x, y, and z directions, computed by our kernel, when its center aligns with various points along the y-axis. Here, the y-coordinates are presented on the x-axis, and the corresponding velocities are plotted on the y-axis. As expected, the y-velocity is the only component exhibiting non-zero values. It effectively propels the robot away from the wall, moving in either a positive or negative direction, depending on the relative position to the wall. Conversely, the x and z velocities remain zero, reflecting the absence of obstacles along these axes. This aligns with our expectations given the presence of the wall obstacle is only along the y direction.

In Figure 4, we observe the variation in the y-position of the fourth joint, both with and without the implementation of the secondary avoidance task. The corresponding plot beneath the image illustrates the diminishing repulsive force as the robot moves further away from the wall obstacle. The rate and function of this decrement can be modulated by altering the dimensions and values within the kernel.

Figure 5 presents a 3D model depicting the final position of the robot, showcasing both scenarios—with the implementation of the joint 4 obstacle avoidance (illustrated in green) and without it (depicted in red).

@@

This paper presented a novel approach to robotic motion planning, combining the application of kernel convolution for obstacle avoidance velocity calculations with an integration of attractive and repulsive field velocities for end-effector trajectory generation and robot joint obstacle avoidance. The method was validated using the kinematic model of a 7 DoF Panda Emika robot, demonstrating its operational versatility and adaptability in varying workspace conditions.

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage{erk}
\usepackage{times}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{xcolor}
% \usepackage{easyReview}
\usepackage[top=22.5mm, bottom=22.5mm, left=22.5mm, right=22.5mm]{geometry}

@@

@@

\todocomment{check that repulsive - avoidance is well changed (READ TEXT)}

@@

\section{Methodology}

@@

The Jacobian matrix is a key element in the control of robotic manipulators. It can also be used for computing the inverse differential kinematics problem. However, for a 7-DOF manipulator, the Jacobian matrix is not square and its inverse does not exist in a traditional sense. In such cases, we resort to the pseudo-inverse of the Jacobian, calculated using the Moore-Penrose method (eq.~\ref{eq:moore_penrose}). The Moore-Penrose pseudo-inverse, denoted as $J^+$, provides an approximation of the inverse for non-square matrices.

Damping, represented by the constant term $k_{damp}$ in the Moore-Penrose pseudo-inverse equation, is used to introduce stability and prevent potential numerical issues when computing the pseudo-inverse of the Jacobian for non-square matrices. It acts as a regularization term to ensure a well-conditioned solution.

@@

The null space (eq.~\ref{eq:Null-Space}) of the Jacobian corresponds to those joint motions which produce no movement in the end-effector. The primary task for our robot is the desired end-effector movement. The secondary task of obstacle avoidance in our research takes advantage of this null-space motion~\cite{lajpah_Petri_2012}. Our method applies an avoidance velocity from detected obstacles in the environment. We calculate this velocities at certain points of interest (POI) that are positioned on the manipulator arm segments. We transform avoidance velocities into the joint space velocities using $J^{\dagger}_{poi}$ jacobian calculated in the specific POI. These velocities we than transform into the null-space using matrix $N$. The robot arm is thus able to carry out the primary task while simultaneously avoiding obstacles by appropriately utilizing its redundant DOF (eq.~\ref{eq:manipulator_kinematics}).

@@

@@

By employing the Hadamard (element-wise) product (eq.~\ref{eq:hadamard}) between the cutout segment of the obstacle grid \(A_d\) and the corresponding 3D convolutional kernel \(K_d\) for direction \(d\), we derive the resultant matrix \(C_d\), which can be expressed as:

@@

Subsequently, the avoidance velocity in the Cartesian coordinate system is obtained by summing all the values in the resulting matrix (eq.~\ref{eq:summation}), which can be represented as:

\begin{equation}
	\dot{x}_{poi} = \sum_{i}\sum_{j}\sum_{k} c_{dijk} 
	\label{eq:summation}
\end{equation}

@@

Our convolutional kernels are essentially discrete differential operators in 3D space, and their design reflects this. These kernels bear a resemblance to the Sobel kernels, which are widely used in computer vision for edge detection and approximation of spatial derivatives. Each kernel has a primary axis aligned with the direction in which the avoidance velocity is being calculated, with the other two axes orthogonal to it. The kernel values along the primary axis exhibit a symmetric pattern (Figure~\ref{Kernel main axis}): the highest value lies in the center of the kernel, with the absolute values decreasing uniformly on either side, with negative polarity on one side. This decrease simulates the derivative-like behavior of the kernel.

@@

Contrary to the design of a classical Sobel operator, our custom kernels don't contain zeroes along the primary axis. Instead, we assign the maximum kernel values along this line. This choice is specifically designed to meet our goal of obstacle avoidance. By doing so, when our points of interest on the robot manipulator coincide with an obstacle - effectively being 'right on top' of the obstacle in the voxel grid - the convolution operation yields a maximal avoidance velocity. This encourages the robot to move away from the obstacle and ensures safer navigation in the presence of potential collisions.

@@

In addition to the kernel values along the primary axis, the values along the orthogonal axes mimic the same pattern (Figure~\ref{Kernel matrix}). These are derived from the primary axis values, decrease proportionally with the radial distance from the primary axis, and crucially, they maintain the same sign as dictated by the derivative along the primary axis. This symmetry in both magnitude and sign around the main axis is a fundamental characteristic of our kernels.

@@

The dimensions of the kernel, both along the primary axis and orthogonal to it, play crucial roles in determining the effective detection range and the nature of the avoidance velocity.

@@

Additionally, we can augment our EE velocity with obstacles avoidance component, which is calculated using the convolutional kernel technique. This enables us to guide the end-effector around obstacles smoothly, ensuring safe and collision-free navigation while dynamically adapting to the environment. By integrating both the attractive and avoidance field contributions, our motion planning approach ensures efficient and effective trajectory generation, enabling the robotic manipulator to reach its desired goal while avoiding potential obstacles in the workspace.

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=1.0\linewidth]{Repulsive-Velocities-Along-the-Y-Axis-wide.eps}
	\caption{Plot depicting avoidance velocities (x, y, z) as calculated by our kernel along the y-axis, showing the robot's adaptive movement away from the wall obstacle.}
	\label{y axis avoidance values}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=1.0\linewidth]{Y-Positions.eps}
	\caption{Comparative visualization of the y-position of the robot's fourth joint with and without the secondary avoidance task, highlighting the impact of the avoidance velocity component.}
	\label{4 th joint plot}
\end{figure}

\begin{figure}[htbp]
	\centering
	\includegraphics[width=1.0\linewidth]{7-DOF_Robot_Positions_Standard_vs_Avoidance.eps}
	\caption{3D model displaying the final position of the robot, contrasting the scenarios with and without the implementation of joint 4 obstacle avoidance.}
	\label{3D model}
\end{figure}

@@

The end-effector's motion was controlled using the attractive field method, serving as the primary task. This was complemented by the avoidance field method implemented as a secondary task, with the specific aim of maneuvering a point on the fourth joint of the robot as far away from the obstacle - in this case, a simple wall - as possible.

@@

Figure~\ref{y axis avoidance values} illustrates the avoidance velocities in the x, y, and z directions, computed by our kernel, when its center aligns with various points along the y-axis. Here, the y-coordinates are presented on the x-axis, and the corresponding velocities are plotted on the y-axis. As expected, the y-velocity is the only component exhibiting non-zero values. It effectively propels the robot away from the wall, moving in either a positive or negative direction, depending on the relative position to the wall. Conversely, the x and z velocities remain zero, reflecting the absence of obstacles along these axes. This aligns with our expectations given the presence of the wall obstacle is only along the y direction.

In Figure~\ref{4 th joint plot}, we observe the variation in the y-position of the fourth joint, both with and without the implementation of the secondary avoidance task. The corresponding plot beneath the image illustrates the diminishing avoidance velocity as the robot moves further away from the wall obstacle. The rate and function of this decrement can be modulated by altering the dimensions and values within the kernel.

Figure~\ref{3D model} presents a 3D model depicting the final position of the robot, showcasing both scenarios—with the implementation of the joint 4 obstacle avoidance (illustrated in green) and without it (depicted in red).

@@

This paper presented a novel approach to robotic motion planning, combining the application of kernel convolution for obstacle avoidance velocity calculations with an integration of attractive and avoidance velocities components for end-effector trajectory generation and robot joint obstacle avoidance. The method was validated using the kinematic model of a 7 DoF Panda Emika robot, demonstrating its operational versatility and adaptability in varying workspace conditions.

@@

\todocomment{add a note about picking up multiple POI}

\todocomment{add note that this is just a part of algorithm development}
